### 2014-05-30 paper.tex
initial paper commit

\documentclass[12pt, titlepage]{article}
% font size could be 10pt (default), 11pt or 12 pt
% paper size coulde be letterpaper (default), legalpaper, executivepaper,
% a4paper, a5paper or b5paper
% side coulde be oneside (default) or twoside 
% columns coulde be onecolumn (default) or twocolumn
% graphics coulde be final (default) or draft 
%
% titlepage coulde be notitlepage (default) or titlepage which 
% makes an extra page for title 
% 
% paper alignment coulde be portrait (default) or landscape 
%
% equations coulde be 
%   default number of the equation on the rigth and equation centered 
%   leqno number on the left and equation centered 
%   fleqn number on the rigth and  equation on the left side
%

\title{Designing Provably Correct Secure Architectures}
\author{Jared Roesch,  \\
    Department of Computer Science, \\
    University of California Santa Barbara. \\
    }

\date{\today} 
% \date{\today} date could be today 
% \date{25.12.00} or be a certain date
% \date{ } or there is no date 
\begin{document}
% Hint: \title{what ever}, \author{who care} and \date{when ever} could stand 
% before or after the \begin{document} command 
% BUT the \maketitle command MUST come AFTER the \begin{document} command! 
\maketitle

\begin{abstract}
Secure hardware is obviously a very active interesting area of research.
People are working hard on desinging architectures that have properties
that are desireable for a vareity of reasons.
How do we reason about \cite{exec-leases}
- restrictions on computation
- static reasoning
- instead of encoding dynamic checks or assuming that

\end{abstract}

%\tableofcontents % create a table of contens

\section{Introduction}

Hardware security is becoming and incresingly important topic both as a research problem and a pragmatic issue effecting 
day to day life. Hardware is multiplying around us and proliferating through all parts of life. We are seeing more and more
devices spring into existance to manage everything from avionics, pace makers, or your thermostat. One of the problems is
that we are giving these devices more and more control over critical systems that govern the safety and lives of humans. At
the same time we also have issues with mobile security, cloud computing, and more.

Traditionally many of these problems have only been thought about at the software level and much of the related work has
been about verifing correctness and security properties about the system running on top.

The problem is that if we want to truly verify correctness properties we must verify each piece of the stack from top to 
bottom.

The missing part of this equation then becomes the verification of hardware. Obviously there has been other work here 
(insert citations) in verifiying simple hardware description, and properties of hardware. What we want to do is introduce
more powerful tools to the hardware developer who actually describes and designs hardware. Unforunately previous approaches
present a DSL in a Theorem Proving environment. There is a endeavor to free the hardware designer to be able to both build
hardware and prove properties.

\section{Background}
- Secure Hardware
- Proving
- DTs
- ect

\section{Secure Architecture Design}

\section{Verified Hardware Construction}

\section{Extensions}

The goal is to take everything that has been described and combine them into something that is uni

\bibliography{paper}
\bibliographystyle{plain}

\end{document}

### 2014-06-01 paper.tex
add paper

@@
 \documentclass[12pt, titlepage]{article}
-% font size could be 10pt (default), 11pt or 12 pt
-% paper size coulde be letterpaper (default), legalpaper, executivepaper,
-% a4paper, a5paper or b5paper
-% side coulde be oneside (default) or twoside 
-% columns coulde be onecolumn (default) or twocolumn
-% graphics coulde be final (default) or draft 
-%
-% titlepage coulde be notitlepage (default) or titlepage which 
-% makes an extra page for title 
-% 
-% paper alignment coulde be portrait (default) or landscape 
-%
-% equations coulde be 
-%   default number of the equation on the rigth and equation centered 
-%   leqno number on the left and equation centered 
-%   fleqn number on the rigth and  equation on the left side
-%
 
-\title{Designing Provably Correct Secure Architectures}
+\input{macros}
+
+\title{Towards a Tool for Verified Hardware Construction}
 \author{Jared Roesch,  \\
@@
 \begin{abstract}
-Secure hardware is obviously a very active interesting area of research.
-People are working hard on desinging architectures that have properties
-that are desireable for a vareity of reasons.
-How do we reason about \cite{exec-leases}
-- restrictions on computation
-- static reasoning
-- instead of encoding dynamic checks or assuming that
-
+Designing secure hardware is an active area of research that has seen much growth in recent years.
+Hardware is now more complicated then ever but even if we state and prove high level security 
+properties about our designs we have no way to demonstrate a correspondence between the thing
+we formally discuss and the thing that we eventually design. Unfortunately the tools used to construct hardware
+have not fundamentally changed in decades. We present a high level overview of the state of the art with regards to these 
+issues, and sketch a plan of how to adapt these techniques in order to build a toolchain for
+the formal verification of hardware synthesis.
 \end{abstract}
@@
 
+\makered{TODO: need to finish putting all my citations in} \\
+
 Hardware security is becoming and incresingly important topic both as a research problem and a pragmatic issue effecting 
 day to day life. Hardware is multiplying around us and proliferating through all parts of life. We are seeing more and more
-devices spring into existance to manage everything from avionics, pace makers, or your thermostat. One of the problems is
+devices spring into existance to manage everything from avionics, pace makers, to your thermostat. One of the problems is
 that we are giving these devices more and more control over critical systems that govern the safety and lives of humans. At
@@
 Traditionally many of these problems have only been thought about at the software level and much of the related work has
-been about verifing correctness and security properties about the system running on top.
+been about verifing correctness and security properties about the systems running on top. Ideally we want to be able an
+run entire stack of critical software; a verified application, running on a verified operating system. This picture
+unforuntately doesn't factor in vulnerabilities at the hardware level. Often times the hardware's security has been
+an afterthought, but there has been multiple recent examples of attacks that leveraged hardware level exploits in order
+to compromise systems(\makered{cite here}).
 
-The problem is that if we want to truly verify correctness properties we must verify each piece of the stack from top to 
-bottom.
+The missing part of this equation then becomes the verification of hardware. Hardware verification is not necessarily a trivial
+problem. Although hardware construction has similarities with design and verification of software, there are 
+key issues that must be solved.
+There has been (\makered{insert citations}) in verifiying simple hardware description, and properties ofhardware.
+(More \makered{limitations here}) The goal then is to build a tool focused on hardware construction that exposes
+more powerful tools to the engineer who designs hardware.
 
-The missing part of this equation then becomes the verification of hardware. Obviously there has been other work here 
-(insert citations) in verifiying simple hardware description, and properties of hardware. What we want to do is introduce
-more powerful tools to the hardware developer who actually describes and designs hardware. Unforunately previous approaches
-present a DSL in a Theorem Proving environment. There is a endeavor to free the hardware designer to be able to both build
+(\makered{incorporate these ideas})
+Unforunately previous approaches present a DSL in a 
+Theorem Proving environment. There is a endeavor to free the hardware designer to be able to both build
 hardware and prove properties.
@@
 \section{Background}
-- Secure Hardware
-- Proving
-- DTs
-- ect
+There has been distinct work in each area of secure computer arhcitectures, and 
+hardware description languages, and theorem proving. We survey the important highlights from each
+that are necessary to understanding our approach.
+\makered{more text here probably}
 
-\section{Secure Architecture Design}
+\subsection{Secure Architecture}
+\begin{itemize}
+    \item Ruby's work, Tim's on execution leases, Appel
+    \item All high level properties that are loosely reasoned about
+    \item want flexibility in design, parametrizable modules (show example from GitHub here)
+    \item Missing rigorous formal analysis
+    \item When they have analysis it suffers from being Pen and Paper, and being abstracted from the actual work.
+\end{itemize}
 
-\section{Verified Hardware Construction}
+In \cite{lee05} they describe “secret-protected” architecture that enables the protection of critical secrets for users in an online environment.
+They argue that their architecture requires only a few extra features to be built into a general purpose microprocessor to protect secrets and related computations.
+They do this with a small amount of extra hardware.
+They are able to decouple user secrets from the devices, avoid symmetric master keys, and avoid doing factory-installed device secrets.
+Their approach can be divided into a few core features the Key Chain, Trusted Software Module, Concealed Execution, Hardware Extensions, Secure I/O, and OS support.
+
+\makered{work some critique in here: The evaluation could use some work.
+There is a written description of both their security and performance evalution, but neither has explicit numbers.
+It seems that although Ruby talks about performance vs.
+secu- rity tradeoffs in her book, she is not completely clear about the ones made here.
+As well they are lacking any sort of formal proof of their security properties.
+This may not be important to the architecture community, but it seems questionable at best to make claims about the security of a system at a high level without actually formally verifiying the properties hold in practice.
+The performance evaluation also seems to be very high level, and ignores important details, for example they dismiss the performance impact on the overall system without presenting precise empirical evaluation of it.
+They do evaluate the instruction load of their hardware crypto but do not compare it aganist anything else in the space.}
+
+In \cite{languagebased} they described ways to add a type system to a small imperative language in order to mitgate the leakage of timing information.
+Ideally these hardware concerns shouldn't leak beyong the hardware level, and they grappel with this in their analysis of the technique.
+
+\makered{put this somewhere: It would be desireable to adapt their technique from \cite{blackboxed} and use our technique to help prove that correctness of the approach.}
+
+\subsection{Hardware Description Languages}
+
+Hardware description languages are software tools for synthesizing hardware from a high level programmatic description.
+A hardware description language is ideally a precise, formal description of an electronic circuit that 
+makes it easy to do automated analysis, simulation, and testing for a circuit. In reality they have
+various weaknesses and failings \subsection{Type Theory}
+Type theory is a foundational framework for mathematics. It was an approach that evolved out of Russel's \cite{Principia}
+as an approach to a consistent foundational system for mathematics. The early 20th century was spent developing type theory
+as a proof system. The key insight was the Curry-Howard Isomorphism, which roughly states that types can be associated with
+propositions and proofs with programs. This insight was adapted and it became apparent that typed variants of the lambda
+calculus can be used as proof systems, the most famous of which is Martin-Lof type theory. You can then view types in
+this typed lambda calculus as various features of constructive logic.
+
+\makered{TODO: clean this section up with typeset examples and such} \\
+
+If we begin with just an implicational fragment of logic we can use the simply typed lambda calculus which only has a
+single type constructor, the function type. This corresponds to implication. We can now view the typing judgment $a : A$
+as statement that a is a proof of A, and we can interepret the typing judgment $f : A -> B$ as a computational implication.
+At the type level it allows us to produce a B given an A, and computationally it gives us a method for transforming a program
+of A into a program of B. If we remember that program's corresond to proofs we now have a computational method for proving.
+
+We can of course
+
+\subsection{Automated Theorem Proving}
+High level talking points: \\
+\begin{itemize}
+    \item dependent types
+    \item equal to a fragment of inutionistic logic
+    \item provides tools for automated proof search
+    \item specification and implementation are one and the same
+    \item can do program extraction
+    \item executable 
+\end{itemize}
+
+Automated Theorem Proving has been an area of active research for many decades. It is apparent that the
+verification of general properties is desireable to both high assurance software engineers, researches, and
+mathmaticians. There have been various approaches to Automated Theorem Proving in the past ACL2, NuPRL, LF,
+LCF, Coq, Agda. The ones based on dependent type theories seem the most promising for various reasons 
+(computational, program extraction, proof carrying code, ect). The cutting edge of type theory has been
+an evolutions of the work done my Per Martin-Lof and others.
+
+Automated Theorem Proving has been applied in many areas from the NTSB, NASA, JPL, and a vareity of other places.
+
+The problem is that very often the specification is seperate from the implementation and suffers the common
+problems of pen and paper proof. In many cases Embedded Domain Specific Languages (EDSLs)
+have been used as a way to gain the power of a proof assitant and also write proof carrying code. We want
+to leverage the ideas here, but ideally built a seperate front-end that exposes the full power of proving,
+but also allows one to provide both an usuable interface that also generates useful proof properties for
+free (\makered{similar to the work done by UCSD PLIDI '14 paper, needs citation}).
+\\
+\makered{NOTE: Work in (For now I'm using a proof environment called Idris for prototyping verified Hardware construction.)}
+
+\subsection{Verified Hardware Construction}
+\begin{itemize}
+    \item Caisson \cite{Li:2011:CHD:1993498.1993512}
+    \item Fe-Si \cite{fesi}
+    \item Cryptol
+    \item Bluespec
+\end{itemize}
 
@@
 
-The goal is to take everything that has been described and combine them into something that is uni
+\begin{itemize}
+    \item consistent design
+    \item demonstrate examples like verification of overflow
+    \item demonstrate bounded input sizes
+\end{itemize}
+
+\begin{itemize}
+    \item consistent design
+    \item syntactic/semantic seperation unlike verilog behavioral/structural
+    \item allow for some properties to be synthesized and proved automaticallly
+    \item verify synthesize and semantics of HDL so that any well formed HDL program has certain properties (ala Coq EDSLs)
+    \item \makered{find citation for UCSD PLDI '14 paper}
+\end{itemize}
+
+Key weaknesses of previous approaches to verification:
+\begin{itemize}
+    \item toy examples
+    \item no automatic property generation/proof satisfaction
+    \item 
+\end{itemize}
+
+\makered{still prototyping the things I want to discuss here}
+\makered{The goal is to take everything that has been described and synthesize the ways in which we can build a tool.}
+
+\section{Conclusion}
+
+\makered{Wrap up, discuss weakness, challenges, ect}
 

### 2014-06-01 paper.tex
more work on dt section

@@
 
-We can of course
+We can of course extend our type system with richer features. We will quickly survey the important additions to the Simply
+Typed Lambda Calculus that are needed to turn it into a dependently typed lambda calculus. The most important feature
+is the idea of Pi Types. Simple type systems can be enriched with many features such as products, sums, so on. There
+have been innovations in type system construction, but depdently typed calculi in the style of the CoC or MLTT are
+the most powerful.
+
+Traditonally we maintain a phase distinction where types and values leave in seperate universes. Depdent types
+are about erasing this distinction. Types can be made more expressive by a series of additions. First if we
+allow quantifers we can now abstract over types with type lambdas. We can then introduce kinds which allow us
+to specify type arities. We can then introduce kind polymorphism, and many other features. Each of these systems become
+more and more complex because they have to maintain constant distinction between each universe of values, kinds, sorts.
+Depdent types simply introduce the idea of a Pi Type of dependent function space. This type system generalization is the 
+idea that we can make everything much simpler by creating a stratfied hierarchy of types, and combining this Pi Types.
+We then can collapse our language into a simple set of pseudo terms, and typing rules \makered{examples can be found}.
+
+The idea of Pi Types can be simply viewed as a generalization of the traditional function type. The generalization
+allows the codomain to depdend on the domain of the function. Given $B : P -> A; \Pi (x : A). B(x)$.
 

### 2014-06-10 paper.tex
finish adding most citations, just need to finish writing

@@
 \section{Introduction}
-
-\makered{TODO: need to finish putting all my citations in} \\
 
@@
 Their approach can be divided into a few core features the Key Chain, Trusted Software Module, Concealed Execution, Hardware Extensions, Secure I/O, and OS support.
+\cite{exec-leases} \cite{PicoCoq2013:POPL}.
+
+\cite{blackboxed} \cite{languagebased} \cite{Chlipala:2007:CTC:1250734.1250742}
 
@@
 
+\cite{HoTTbook} \cite{Pierce:TypeSystems}
+
 \subsection{Automated Theorem Proving}
@@
 but also allows one to provide both an usuable interface that also generates useful proof properties for
-free (\makered{similar to the work done by UCSD PLIDI '14 paper, needs citation}).
+free \cite{Ricketts:2014}. \cite{chlipala2011certified} \cite{Pierce:SF}
 \\
@@
 \subsection{Verified Hardware Construction}
-\begin{itemize}
-    \item Caisson \cite{Li:2011:CHD:1993498.1993512}
-    \item Fe-Si \cite{fesi}
-    \item Cryptol
-    \item Bluespec
-\end{itemize}
+There has been work on creating tools for doing designing secure hardware such as Caisson and SAFE but in both cases the
+compiler assumes and enforces a standardized security policy. \makered{elaborate} \cite{Li:2011:CHD:1993498.1993512}
+There also has been work on specialized systems like Cryptol that make simplifying assumptions allowing certain gaurentees
+and properties to hold. There has been some related work in building higher level design languages such as Lava, and Bluespec
+that allow the hardware designer to abstract over properties such as explicit timing information. The most closely related
+work has been Fe-Si \cite{fesi} \cite{Brady_constructingcorrect} \cite{cryptol2010}
 
@@
     \item verify synthesize and semantics of HDL so that any well formed HDL program has certain properties (ala Coq EDSLs)
-    \item \makered{find citation for UCSD PLDI '14 paper}
+    \item \makered{find citation for UCSD PLDI '14 paper} \cite{Ricketts:2014}
 \end{itemize}
@@
 
-\makered{Wrap up, discuss weakness, challenges, ect}
+There has been a lot of work poured into constructing architectures that exhibit desired security properites. Unfortunately there hasn't been
+much work on the implementation and verification of these architectures so although on paper we are able to reason about the behavior of the
+system we don't have many guarentees about the behavior of our actual implementation. This is a failing of the tools used by hardware
+designers. We present a blueprint of how to leverage work done in the programming languages community and leverage automated theorem provers
+and dependent types as tools for specifying and verifying properties of hardware designs. We leave the more diffcult problem of designing and
+verifying hardware designs to future work, the first goal is to build a tool that is powerful and robust enough to actually port real designs
+to it.
 

### 2014-06-10 paper.tex
nearing final draft

--- 2014-06-01 paper.tex
@@
 Designing secure hardware is an active area of research that has seen much growth in recent years.
-Hardware is now more complicated then ever but even if we state and prove high level security 
-properties about our designs we have no way to demonstrate a correspondence between the thing
-we formally discuss and the thing that we eventually design. Unfortunately the tools used to construct hardware
-have not fundamentally changed in decades. We present a high level overview of the state of the art with regards to these 
-issues, and sketch a plan of how to adapt these techniques in order to build a toolchain for
+Hardware is now more complicated then ever but even if we state and prove; high level security 
+properties about our designs, we have no way to demonstrate a correspondence between the model
+we formally discuss and the implementation that we eventually design. In the face of changing
+demands the tools used to construct hardware have not fundamentally changed in decades. 
+We present a high level overview of the state of the art with regards to these 
+issues, and sketch a plan of how to adapt these techniques in order to build a tool chain for
 the formal verification of hardware synthesis.
@@
 
-\makered{TODO: need to finish putting all my citations in} \\
+Hardware security is becoming and increasingly important topic both as a research problem and one that has an effect on
+day to day life. Hardware is multiplying around us and proliferating through all parts of life. We are seeing more and more
+devices spring into existence to manage everything from avionics, pace makers, to your thermostat. An issue then
+becomes, that as we give these devices more and more control, we are more and more interested in moderating their control. This is
+because control of these devices isn't just the device or its contents as it has historically been, but access to sensitive personal
+control, for example an exploited iPhone may have greater implications when it can open your doors, or command household electronics.
+Households could be made vulnerable to "malfunction" whether malicious or not. As well there are issues to be had with mobile security, as 
+smart phones play a more and more important role in in daily life. It goes hand in hand with cloud computing, which is playing out as a nice
+dual to the rise of smart phones, more and more computation has been moved centrally to the "cloud" and it introduces central points of failure
+if they are compromised. No matter the platform it is necessary to begin ensuring the integrity of our software from the top down. Even if we
+introduce notions of verified software it is meaningless in the face of exploitable hardware.
 
-Hardware security is becoming and incresingly important topic both as a research problem and a pragmatic issue effecting 
-day to day life. Hardware is multiplying around us and proliferating through all parts of life. We are seeing more and more
-devices spring into existance to manage everything from avionics, pace makers, to your thermostat. One of the problems is
-that we are giving these devices more and more control over critical systems that govern the safety and lives of humans. At
-the same time we also have issues with mobile security, cloud computing, and more.
-
-Traditionally many of these problems have only been thought about at the software level and much of the related work has
-been about verifing correctness and security properties about the systems running on top. Ideally we want to be able an
-run entire stack of critical software; a verified application, running on a verified operating system. This picture
-unforuntately doesn't factor in vulnerabilities at the hardware level. Often times the hardware's security has been
-an afterthought, but there has been multiple recent examples of attacks that leveraged hardware level exploits in order
-to compromise systems(\makered{cite here}).
+Traditionally the specification and verification of security properties has only been done at an architectural 
+level, by either design or by a baked in policy, and enforcement . A lot of work has been done about
+properties been about verifying correctness and security properties about the systems running on top. The ideal presented 
+in older works was to be able to run an entire stack of critical software; a verified application, running on a verified operating system. 
+This picture unfortunately doesn't factor in vulnerabilities at the hardware level. Often times the hardware's security has been
+an afterthought, but there has been many recent examples of attacks that leveraged hardware level exploits in order
+to compromise systems.
 
@@
 problem. Although hardware construction has similarities with design and verification of software, there are 
-key issues that must be solved.
-There has been (\makered{insert citations}) in verifiying simple hardware description, and properties ofhardware.
-(More \makered{limitations here}) The goal then is to build a tool focused on hardware construction that exposes
-more powerful tools to the engineer who designs hardware.
+key issues that must be solved. There has been success in verifying simple hardware description, 
+and properties of the designs. The goal is to take the kernel of an idea and flesh it out into
+a tool focused on hardware construction that exposes very powerful tools to the engineer who designs 
+hardware.
 
-(\makered{incorporate these ideas})
-Unforunately previous approaches present a DSL in a 
-Theorem Proving environment. There is a endeavor to free the hardware designer to be able to both build
-hardware and prove properties.
+Unfortunately the state-of-the-art \cite{fesi} only presents a limited domain specific language
+in a theorem proving environment, providing very little support to to a hardware designer who wants to be 
+able to write designs and then prove properties. We believe that it is possible to do better providing a 
+pragmatic and useful tool that allows us to leverage the power of some properties for free, as well
+as the ability to add more via the proof system, enabling hardware design.
 
 \section{Background}
-There has been distinct work in each area of secure computer arhcitectures, and 
+There has been distinct work in each area of secure computer architectures, and 
 hardware description languages, and theorem proving. We survey the important highlights from each
 that are necessary to understanding our approach.
-\makered{more text here probably}
 
 \subsection{Secure Architecture}
-\begin{itemize}
-    \item Ruby's work, Tim's on execution leases, Appel
-    \item All high level properties that are loosely reasoned about
-    \item want flexibility in design, parametrizable modules (show example from GitHub here)
-    \item Missing rigorous formal analysis
-    \item When they have analysis it suffers from being Pen and Paper, and being abstracted from the actual work.
-\end{itemize}
 
@@
 Their approach can be divided into a few core features the Key Chain, Trusted Software Module, Concealed Execution, Hardware Extensions, Secure I/O, and OS support.
+\cite{PicoCoq2013:POPL}.
+It seems that although Lee talks about performance vs.
+Security trade offs in her book, though their work is not clear about which trade offs they made.
+They also lack formal proof of their security properties.
+This may not be important to the architecture community, but it seems questionable at best to make claims about the security of a system at a high level without actually formally verifying the properties hold in practice.
 
-\makered{work some critique in here: The evaluation could use some work.
-There is a written description of both their security and performance evalution, but neither has explicit numbers.
-It seems that although Ruby talks about performance vs.
-secu- rity tradeoffs in her book, she is not completely clear about the ones made here.
-As well they are lacking any sort of formal proof of their security properties.
-This may not be important to the architecture community, but it seems questionable at best to make claims about the security of a system at a high level without actually formally verifiying the properties hold in practice.
-The performance evaluation also seems to be very high level, and ignores important details, for example they dismiss the performance impact on the overall system without presenting precise empirical evaluation of it.
-They do evaluate the instruction load of their hardware crypto but do not compare it aganist anything else in the space.}
+In \cite{blackboxed} \cite{languagebased} they discuss mechanisms for architectural and language support for mitigating the
+leakage of information through side channels. They take the approach of inserting time stops in order to prevent implicit
+leaks from control flow structures.
 
-In \cite{languagebased} they described ways to add a type system to a small imperative language in order to mitgate the leakage of timing information.
-Ideally these hardware concerns shouldn't leak beyong the hardware level, and they grappel with this in their analysis of the technique.
+In \cite{languagebased} they described ways to add a type system to a small imperative language in order to mitigate the leakage of timing information.
+Ideally these hardware concerns shouldn't leak beyond the hardware level, and they grapple with this in their analysis of the technique.
 
-\makered{put this somewhere: It would be desireable to adapt their technique from \cite{blackboxed} and use our technique to help prove that correctness of the approach.}
+In \cite{exec-leases} they also detail an information flow architecture base upon their GLIFT technique. The critical idea of the work is that
+we can provide secure leases that allow for a caller to bound the privileges both in terms of information and time for the lessee.
+
+Having identified some of the weaknesses in the implementations of secure architectures we would 
+ideally like to compliment them by providing a tool that enables the following criteria:
+\begin{itemize}
+    \item state high and low level security properties
+    \item reason about abstract specification
+    \item correspondence between specification and hardware
+    \item missing rigorous formal analysis
+\end{itemize}
 
@@
 A hardware description language is ideally a precise, formal description of an electronic circuit that 
-makes it easy to do automated analysis, simulation, and testing for a circuit. In reality they have
-various weaknesses and failings \subsection{Type Theory}
-Type theory is a foundational framework for mathematics. It was an approach that evolved out of Russel's \cite{Principia}
-as an approach to a consistent foundational system for mathematics. The early 20th century was spent developing type theory
+makes it easy to do automated analysis, simulation, and testing of a circuit. In reality they have
+various weaknesses and failings that mostly result from the weaknesses introduced by current tooling. 
+We would ideally like to make it easier to state precise specifications of hardware, and verify that
+the implementations matches said specification.
+
+\subsection{Type Theory}
+Type theory is a foundational framework for mathematics. It was an approach that evolved out of Russel's work on setting
+mathematics on a consistent foundational system for mathematics. The early 20th century was spent developing type theory
 as a proof system. The key insight was the Curry-Howard Isomorphism, which roughly states that types can be associated with
 propositions and proofs with programs. This insight was adapted and it became apparent that typed variants of the lambda
-calculus can be used as proof systems, the most famous of which is Martin-Lof type theory. You can then view types in
+calculus can be used as proof systems, the most widely used of which is Martin-Lof type theory. You can then view types in
 this typed lambda calculus as various features of constructive logic.
-
-\makered{TODO: clean this section up with typeset examples and such} \\
 
@@
 single type constructor, the function type. This corresponds to implication. We can now view the typing judgment $a : A$
-as statement that a is a proof of A, and we can interepret the typing judgment $f : A -> B$ as a computational implication.
+as statement that a is a proof of A, and we can interpret the typing judgment $f : A -> B$ as a computational implication.
 At the type level it allows us to produce a B given an A, and computationally it gives us a method for transforming a program
-of A into a program of B. If we remember that program's corresond to proofs we now have a computational method for proving.
+of A into a program of B. If we remember that program's correspond to proofs we now have a computational method for proving.
 
-We can of course
+We can of course extend our type system with richer features. We will quickly survey important additions to the Simply
+Typed Lambda Calculus that are needed to turn it into a dependently typed lambda calculus. The key idea is the erasure
+of phase distinction via the inclusion of a stratified type hierarchy, and Pi types.
 
-\subsection{Automated Theorem Proving}
-High level talking points: \\
+Traditionally we maintain a phase distinction where types and values live in separate universes. Dependent types
+are about easing this distinction. Type systems can usually be made more expressive by a series of additions.
+The simplest of which is the introduction of quantifiers which allow us to abstract over types with type lambdas. 
+We can then introduce kinds (the types of types), and kind polymorphism which allow us to specify type shapes or (type arities).
+We can continue to enrich our kind and type system, eventually reflecting types as kinds, and values as types, 
+but these systems are actually more complicated then a dependently typed calculi. Each of these refines become
+more and more complex because it requires the language to maintain a discipline between values, kinds, sorts.
+Dependent types simply introduce the idea of a Pi Type of dependent function space. This type system generalization is the 
+idea that we can make everything much simpler by creating a stratified hierarchy of types, and adding Pi and Inductive types.
+We then can collapse our language into a simple set of pseudo terms, and typing rules which are easy to verify for correctness.
+
+We can then ignore all the complexity introduced by types systems such as System F and generalizations of it. Instead we can 
+focus on a simple language. Usually a dependently typed core only requires a few forms abstraction, variables, function application,
+and Pi Types. This is simply the lambda calculus that everyone knows enriched with Pi types. It is also useful to extend the calculus
+with inductive types, or recursive data types that may be parametrized by the values they are constructed with.
+
+Pi Types can be simply viewed as a generalization of the traditional function type. The generalization
+allows the type of codomain to depend on value of the domain of the function. Given $B : P -> A; \Pi (x : A). B(x)$. This
+simple extension actually gives rise to a lot of power and will allow us to encode much of mathematics in type theory.
+
+This is of course a very brief overview and more information can be found in \cite{Pierce:TypeSystems} \cite{martinlof} and \cite{HoTTbook}.
+
+\subsection{Theorem Proving}
+
+Automated Theorem Proving has been an area of active research for many decades. It is apparent that the
+verification of general properties is desirable to both high assurance software engineers, researches, and
+mathematicians. There have been various approaches to Automated Theorem Proving in the past ACL2, NuPRL, LF,
+LCF, Coq, Agda. The ones based on dependent type theories seem the most promising for various reasons 
+(computational, program extraction, proof carrying code, ect). Automated Theorem Proving has been applied in
+many areas from the NTSB, NASA, JPL, and a variety of other places.
+
+The current best way to leverage the proving power of Type Theory are Automated or Interactive Theorem Proving
+environments. These allow for the automatic satisfaction of some properties, and interactive proving of others.
+The state-of-the-art theorem proving environments in use for software verification are Coq, Agda, and Idris 
+all of which provide the benefits of:
 \begin{itemize}
-    \item dependent types
-    \item equal to a fragment of inutionistic logic
+    \item dependent types (equivalent to a fragment of isolationistic logic)
     \item provides tools for automated proof search
     \item specification and implementation are one and the same
-    \item can do program extraction
-    \item executable 
+    \item executable "proofs" (i.e program extraction)
 \end{itemize}
 
+These are all based on versions of intensional type theory. Coq is base on
+on Coquand and Huet's Theory of Constructions, and Agda is an evolution 
+Martin-Lof Type Theory \cite{martinlof}.
+
 Automated Theorem Proving has been an area of active research for many decades. It is apparent that the
-verification of general properties is desireable to both high assurance software engineers, researches, and
-mathmaticians. There have been various approaches to Automated Theorem Proving in the past ACL2, NuPRL, LF,
+verification of general properties is desirable to both high assurance software engineers, researches, and
+mathematicians. There have been various approaches to Automated Theorem Proving in the past ACL2, NuPRL, LF,
 LCF, Coq, Agda. The ones based on dependent type theories seem the most promising for various reasons 
-(computational, program extraction, proof carrying code, ect). The cutting edge of type theory has been
-an evolutions of the work done my Per Martin-Lof and others.
+(computational, program extraction, proof carrying code, ect). Th current state-of-art is either based
+on Coquand and Huet's Theory of Constructions or the work done on Martin-Lof Type Theory \cite{martinlof}.
+Automated Theorem Proving has been applied in many areas from the NTSB, NASA, JPL, and a variety of other places.
 
-Automated Theorem Proving has been applied in many areas from the NTSB, NASA, JPL, and a vareity of other places.
+It appears that ideas from ATP could be very useful in the domain of hardware specification. Most
+specification is performed separately from the implementation and suffers the common
+problems of pen and paper proof. In many cases Embedded Domain Specific Languages (EDSLs)
+have been used as a way to gain the power of a proof assistant and also write proof carrying code\cite{Ricketts:2014}\cite{fesi}. We want
+to leverage the ideas here, but ideally built a separate front-end that exposes the full power of proving,
+but also allows one to provide both an usable interface that also generates useful proof properties for
+free \cite{Ricketts:2014}. \cite{chlipala2011certified} \cite{Pierce:SF}
 
-The problem is that very often the specification is seperate from the implementation and suffers the common
-problems of pen and paper proof. In many cases Embedded Domain Specific Languages (EDSLs)
-have been used as a way to gain the power of a proof assitant and also write proof carrying code. We want
-to leverage the ideas here, but ideally built a seperate front-end that exposes the full power of proving,
-but also allows one to provide both an usuable interface that also generates useful proof properties for
-free (\makered{similar to the work done by UCSD PLIDI '14 paper, needs citation}).
-\\
-\makered{NOTE: Work in (For now I'm using a proof environment called Idris for prototyping verified Hardware construction.)}
+There is a prototype implementation of ideas discussed in the final section available. It is currently
+written in Idris one of the more practical dependently typed programming language.
 
 \subsection{Verified Hardware Construction}
-\begin{itemize}
-    \item Caisson \cite{Li:2011:CHD:1993498.1993512}
-    \item Fe-Si \cite{fesi}
-    \item Cryptol
-    \item Bluespec
-\end{itemize}
+
+There has been work on creating tools for aiding in the design of secure hardware such as Caisson which presents a HDL 
+language that has a preselected static information flow policy. This provides a information flow policy by baking in 
+a static analysis to the compiler. We would like to be able to easily experiment with different different security 
+properties and be able generally state arbitrary security properties of the hardware. We can easily discharge these
+proof assumptions by a mixture of proof automation and interactive proving. SAFE takes a similar approach to the
+compiler assumes and enforces a standardized security policy. This is also the case with Caisson a gate level 
+approach that is baked in to the HDL compiler \cite{Li:2011:CHD:1993498.1993512}.
+
+There also has been work on specialized systems like Cryptol that make simplifying assumptions allowing certain guarantees
+and properties to hold \cite{cryptol2010}. Cryptol is able to use tools like SMT solvers and simple sized type system in
+order to prove properties about cryptography implementations. They are able to make these simplifications be identifying 
+ways in which traditional cryptography hardware is constructed and designing their language accordingly.
+
+ There has been some related work in building higher level design languages such as Lava, and Bluespec
+that allow the hardware designer to abstract over properties such as explicit timing information. The most closely related
+work has been Fe-Si \cite{fesi} which can viewed as a deterministic version of BlueSpec. The key insights to take away from
+the Fe-Si work is the description of how to use the meta languages type system for great effect, as well as inight into
+the verificatin of HDL compiler passes. Adam Chilpala has done a lot work on describing novel ways in which to verify different
+aspects of programming languages \cite{Chlipala:2007:CTC:1250734.1250742}\cite{chlipala2011certified}
+
+There also has been some work from the dependently typed programming language community sketching ways in which dependent types
+can be made useful in hardware design\cite{Brady_constructingcorrect}. The authors only scratch the surface of the issues that
+occur in real hardware design and take a tutorial like approach to describing how to leverage dependent types to assist in 
+design correctness, demonstrating very simple combinational designs like adders.
 
@@
 
+Our general goals are to leverage ideas from program analysis, type systems and
+apply them to a hardware construction language. We motivate our desire to replace existing HDLs for the
+reasons we listed in the previous section. Primarily it is about pursuing a more consistent design, that
+allows us to have both concise hardware descriptions, and write proofs about said specifications. We
+also want to equip the surface language with the ability to demand certain properties and proofs for
+free. For example we want to provide guarantees of properties such as overflow, prop wiring by construction,
+and so on, but leave the ability to establish higher level proof properties like the information
+flow expressed by default in Caisson's compiler \cite{Li:2011:CHD:1993498.1993512}.
+
 \begin{itemize}
-    \item consistent design
-    \item demonstrate examples like verification of overflow
-    \item demonstrate bounded input sizes
+    \item allow for some properties to be synthesized and proved automatically
+    \item verify synthesize and semantics of HDL so that any well formed HDL program has certain properties similar to \cite{Ricketts:2014}
 \end{itemize}
 
-\begin{itemize}
-    \item consistent design
-    \item syntactic/semantic seperation unlike verilog behavioral/structural
-    \item allow for some properties to be synthesized and proved automaticallly
-    \item verify synthesize and semantics of HDL so that any well formed HDL program has certain properties (ala Coq EDSLs)
-    \item \makered{find citation for UCSD PLDI '14 paper}
-\end{itemize}
+Most previous approaches have not been practical enough to use for building real designs. Fe-Si is as close to
+complete but is still unfortunately just a toy. The DSL approach has important limitations and requires
+programming in an awkward monadic style inside of the theorem proof system Coq. Ideally we would like to take the 
+simple core described in the Fe-Si work, but extend it in a few key ways. First we would like to add a surface language
+that will give us the ability to add more complicated features than the limited Fe-Si. We will also like to make modules first class
+entities that express information about the functionality of the circuit as well as
 
-Key weaknesses of previous approaches to verification:
-\begin{itemize}
-    \item toy examples
-    \item no automatic property generation/proof satisfaction
-    \item 
-\end{itemize}
-
-\makered{still prototyping the things I want to discuss here}
-\makered{The goal is to take everything that has been described and synthesize the ways in which we can build a tool.}
+We have experimented with a few different designs, and am still considering which is the best way forward.
+So far we have not made a decision on which is best, but have noticed some important trade offs in the process
+of implementing
 
@@
 
-\makered{Wrap up, discuss weakness, challenges, ect}
+There has been a lot of work poured into constructing architectures that exhibit desired security properties. Unfortunately there hasn't been
+much work on the implementation and verification of these architectures so although on paper we are able to reason about the behavior of the
+system we don't have many guarantees about the behavior of our actual implementation. This is a failing of the tools used by hardware
+designers. We present a blueprint of how to leverage work done in the programming languages community and leverage automated theorem provers
+and dependent types as tools for specifying and verifying properties of hardware designs. We leave the more difficult problem of designing and
+verifying hardware designs to future work, the first goal is to build a tool that is powerful and robust enough to actually port real designs
+to it.
 

### 2014-06-10 paper.tex
add more english that appeared to be missing

@@
 leakage of information through side channels. They take the approach of inserting time stops in order to prevent implicit
-leaks from control flow structures.
-
-In \cite{languagebased} they described ways to add a type system to a small imperative language in order to mitigate the leakage of timing information.
-Ideally these hardware concerns shouldn't leak beyond the hardware level, and they grapple with this in their analysis of the technique.
+leaks from control flow structures. In \cite{languagebased} they described ways to add a type system to a small imperative language in 
+order to mitigate the leakage of timing information. Ideally these concerns shouldn't leak beyond the hardware level, and they grapple with this in their analysis of the technique.
 
@@
 There is a prototype implementation of ideas discussed in the final section available. It is currently
-written in Idris one of the more practical dependently typed programming language.
+written in Idris (which is one of the more practical dependently typed programming languages).
 
@@
 that will give us the ability to add more complicated features than the limited Fe-Si. We will also like to make modules first class
-entities that express information about the functionality of the circuit as well as
+entities that express information about the functionality of the circuit as well so that we can use such information when proving
+properties about designs.
 
@@
 So far we have not made a decision on which is best, but have noticed some important trade offs in the process
-of implementing
+of implementing prototypes. Many of the naive designs don't work because they lose precious typing information
+that is essential for proving properties about the design. It has become apparent that many of the decisions made
+by Adam Chilpalia's work are important. We will use the same approach of using a denontational approach, and relying
+on properties of the metalanguage such as strong normalization, and its type system to assure that our designs are
+correct by construction. We could of course elaborate our language to a core calculus and implement our own checker
+but it is important to not duplicate effort early in the research process, and should piggy back on existing work.
+For now we are translating directly to Idris source language, but it exposes its core calculus and checker as a
+library we can eventually use if we want to piggyback on their implementation effort. We also want to model his
+approach to slowly adding and proving the correctness of optimizations beyond the simple ones performed in \cite{fesi}.
+
+One of the long term goals is a verified framework for abstract interpretation a powerful analysis technique, but for
+now we will continue on in the style of Chilpalia.One of the big departures from FeSi is that we want to provide a back door 
+into the full theorem proving environment which will allow for designers or verifiers to add more complex theorems, and maintain 
+the power provided by a full proof assitant such as Coq.
 
